---
title: "Dev Submission Assignment - 1"
output:
  html_document:
    df_print: paged
---
#Q1 Dataset with a mix of quantitative and qualitative (categorical) variables. Taken from - (https://www.kaggle.com/borapajo/food-choices#food_coded.csv)


Q2: Reading and importing dataset into R

```{r}
data <-read.csv("C:/Users/BrandRely/Documents/Datasets/food.csv")
head(data)
```

Q3: Descriptive statistics for a selection of quantitative and categorical variables

```{r}
summary(data)

#Quantative variable Descriptive Statistics
mean(data$GPA) 
sd(data$GPA)
median(data$GPA)
var(data$GPA)

#Categorical variable Descriptive Statistics 
table(data$Gender)
table(data$drink)
```

Q4: Transforming variable "Transform" as sum of two colums - "tortilla_calories" + "calories_chicken" & taking log of newly created variable

```{r}
data$chic_and_tort <- data$tortilla_calories + data$calories_chicken
data$logtrnas <- log10(data$chic_and_tort) 
data$logtrnas
colnames(data)
```

Q5: Plot at least one quantitative variable, and one scatterplot

```{r}
# Histogram

hist(data$GPA, main ="Frequency of GPA", xlab = "GPA", ylab = "Frequency", border="red", col="blue")

#Scatterplot

plot(data$GPA, data$chic_and_tort,  main='Barplot', xlab = "GPA", ylab = "Chic & Tort", col='yellow')
```

