---
title: "Dev Submission Assignment 3"
output: html_document
---

Step 1: Loading important libraries

```{r}

library(ggplot2)
library(naivebayes)
library(pROC)
library(gmodels)
library(caret)
library(e1071)
library(dplyr)
library(psych)
```

Step 2: Reading the file & manipulating data as required

```{r}
Original_Data <- read.csv("C:/Users/BrandRely/Documents/Datasets/FlightDelays.csv")

#Selecting only columns needed

New_Data <- Original_Data[,c(1,2,4,8,10,13)]

New_Data$DAY_WEEK <- as.factor(New_Data$DAY_WEEK)
New_Data$CRS_DEP_TIME <- as.factor(New_Data$CRS_DEP_TIME)
summary(New_Data)
```

Step 3: Partitioning data into 60% Training and 40% Validation (Question - 1)

```{r}
set.seed(123)
Train_Index <- createDataPartition(New_Data$Flight.Status, p=0.6, list=FALSE)
Train_Data <- New_Data[Train_Index,]
Validation_Data <- New_Data[-Train_Index,]
```


Step 4: Running Naive Bayes Model to predict Flight Delay (Question - 2)

```{r}
naive_model_1 <- naive_bayes(Flight.Status ~ ., data =Train_Data, usekernel = T)
naive_model_1
```

Step 5: Showing Flights that were delayed and on-time at each of the three airports (Question - 3)

```{r}
#Destination Table
destination_table <-  Train_Data %>% count(DEST, Flight.Status) %>% mutate(proportion_table = prop.table(n))
destination_table

#Forming Count Table
Total_Count <- table(Train_Data$Flight.Status,Train_Data$DEST)
Total_Count

#Forming Proportion Table
Proportion_Table <- prop.table(Total_Count)
Proportion_Table
```

Step 6: Confusion matrix & ROC for the Validation Data (Question - 4)

```{r}
#Training Set

Predicted_Test_Label_1 <- predict(naive_model_1,Train_Data)
confusionMatrix(table(Predicted_Test_Label_1,Train_Data$Flight.Status))

#Validation Set

Predicted_Test_Label_2 <- predict(naive_model_1,Validation_Data)
confusionMatrix(table(Predicted_Test_Label_2,Validation_Data$Flight.Status))

#Forming Cross-Table
CrossTable(x=Validation_Data$Flight.Status,y=Predicted_Test_Label_2, prop.chisq = FALSE) 

naive_model_2 <- naiveBayes(Flight.Status ~.,data = Train_Data)
Predicted_Test_Label_3 <- predict(naive_model_2,Validation_Data, type = "raw")

head(Predicted_Test_Label_3)

#Creating the ROC

roc(Validation_Data$Flight.Status, Predicted_Test_Label_3[,1])
plot.roc(Validation_Data$Flight.Status, Predicted_Test_Label_3[,1])
```